A Temporal Logic Based Framework for Intrusion Detection

### 2023-09-16 Paper/paper.tex
update

\section{MONID: Structure analysis}
MONID is a prototype which can detect intrusions either online or offline. System-level event data is delivered to MONID server either directly from relevant log files (offline) or online via application code that has been properly instrumented. In the online mode, MONID runs as a server that receives streams of events from various sources and to generate these events, the different logging modules are instrumented so that they filter and send relevant events to the server.
To create a single event trace, the server combines events from numerous sources according to timestamps and preprocesses them into an abstract intermediate form. Note that sometimes it is necessary to watch events from multiple sources in order to detect certain attacks; then, MONID monitor checks this event trail against a predetermined criteria, and if the specification is broken, an intrusion alarm is raised. While, in the offline mode, MONID reads various log files and sends an event corresponding each log entry to the server. The server then processes the event stream as before to detect intrusion.

@@

Some evaluations were also have been done in~\cite{naldurg2004temporal}, using the DARPA Intrusion Detection Evaluation dataset~\cite{mchugh2000testing}. The experiment was done using offline mode, against some of the attacks presented in Section~\ref{attacks}. More specifically, against the smurf, port-sweep and password-guessing attacks; the experiments detected 5 password-guess attack and 2 port-sweep attack in the logs and the performance overheads for monitoring the Port-Sweep and Password-Attacks are reported in Figure~\ref{graph1} and~\ref{graph2}, respectively. The X-axis in both graphs shows the number of events we are monitoring, as they are obtained from our logs. Each data point is the average overhead calculated for intervals of 10 and 1000 events respectively. The Y-axis plots the ratio between the time spent by the monitor vs. the time between the generation of the events in the actual log. As long as this ratio is less than 1, our monitoring is feasible. The results, show that the average overhead, is around $0.009$ for Port-Sweep attacks and $0.016$ for Password-Guessing attacks.
Although the used hardware is now outdated and obsolete, the overhead was rather modest, thus this method should still be theoretically helpful today.

@@

\section{Example of attacks}\label{attacks}
We give some examples of how EAGLE can be utilized to build formulas that correlate to desired execution trace characteristics of a system under observation in this section. In this context, a trace $\sigma$ that violates this specification is referred to as an intrusion or an attack; to explain how the framework may be used, we use examples from actual attacks illustrated in~\cite{naldurg2004temporal}. These illustrations demonstrate the expressive power of EAGLE formalism as well as its numerous features. In all the following examples, the rule Always is the one defined in Section~\ref{sec:ltl}.

\textbf{Smurf attacks.}
A smurf attack~\cite{zargar2009identification} is a type of Distributed Denial of Service (DDoS) attack that targets computer networks; in this scenario, the attacker sends a large volume of Internet Control Message Protocol (ICMP) echo request packets (commonly known as "ping" packets) to a network's broadcast address. This broadcast address is the address that reaches all devices on the targeted network. The attacker spoofs the source IP address of these ICMP packets to make it appear as if they are originating from the victim's IP address and since the broadcast address sends the request to all devices on the network, all devices will respond with ICMP echo replies to the victim's IP address. This results in a flood of traffic overwhelming the victim's network and causing it to become unreachable or severely degraded in performance. To detect this kind of attacks, we need to work on network logs; we will use a network tool called \textit{tcpdump}. The monitored formula that describes the absence of the attack is the following:
\begin{align*}
    & \underline{\text{max}}\ \mathtt{SmurfAttack()} = (type = \text{"ICMP"})\land isBroadcast(ip) \\
    & \underline{\text{mon}}\ \mathtt{SmurfSafety} =\ \mathtt{Always}(\neg\ \mathtt{SmurfAttack()}) 
\end{align*}
as we can see, the \textit{type} field of a record schema describes the type of network packet, while the \textit{ip} field is the destination of the packet. The $\mathtt{SmurfAttack}$ rule determines whether the packet is a ping packet and whether the source's IP address is a broadcast ip, i.e. all the host bits are set to 1. Thus, \textit{"Always there is no attack"} can be used to describe how the system behaves well in response to this attack.

\textbf{Cookie-stealing attacks.}
A cookie is a technique for tracking sessions that a web application server sends to a web client and uses to keep track of client-specific session data. These cookies that serve as authentication tokens are automatically included in requests by clients and in this scenario we assume that a session is exclusively identified by the client's IP address. When a malicious user uses an old cookie provided to a different IP address to take over a session, it is considered an attack.In order to monitor this attack we need to look at a web-server (application-level) log that contains a record of all sessions that the server participates in, along with session-specific state information.
According to the formula below, a specific cookie must always be utilized by the same IP address.
\begin{align*}
    & \underline{\text{min}}\ \mathtt{Hijack}(\underline{\text{string}}\ c,\underline{\text{string}}\ i) = ((name=c)\land\neg(ip=i))\lor\odot\ \mathtt{Hijack}(c,i) \\
    & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\neg\mathtt{Hijack}(name,ip)) 
\end{align*}
A trace that violates this formula therefore encodes a cookie-stealing attack. The rule $\mathtt{Hijack}$ checks wheter it occurs a situation in which we have the same coookie $c$ associated with two different ip $i$; if it's true then it means that a cookie has been stolen and reused by an attacker. With $\mathtt{CookieSafe}$ we want to check that it never occurs a situation just as the one described.

\textbf{Multi-domain Buffer Overflows attacks.}
According to~\cite{piromsopa2006buffer}, a buffer-overflow attack is an attack that (possibly implicitly) uses memory-manipulating operations to overflow a buffer which results in the modification of an address to point to malicious or unexpected code. In this case, we will integrate data from various log sources; to be more exact, we will examine both network and web server access logs to determine when a server buffer overflow attack has been performed. We search through network packets for binary data; then web server's access logs are examined to see if a corresponding event, in which the web server correctly terminates the connection after receiving some binary data, can be located. If no matching log record is discovered within a predetermined timeout, it means that the buffer overflow attack was successful. The following formulas captures this scenario:
\begin{align*} 
    & \underline{\text{min}}\ \mathtt{EventuallyClosed}(\underline{\text{long}}\ t,\underline{\text{long}}\ d,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2) = (time-t<d)\\
    & \quad\quad\land ((ip1=i1\land ip2=i2\land log=web\land type=closed)\lor \bigcirc\mathtt{EventuallyClosed}(t,d,i1,i2) ) \\
    & \underline{\text{mon}}\ \mathtt{BufferSafe} =\ \mathtt{Always}((log=network\land type=binary) \\
    & \quad\quad\rightarrow\mathtt{EventuallyClosed}(time,100,ip1,ip2))
\end{align*}
In each record we have the $log$ field which indicates the source of the log, the $time$ represents the istant in which the event occurs, $ip1$ represents the source address, $ip2$ represents the destination IP address and finally the $type$ field can be $binary, closed,\ldots$ We assume that a connection is identified by the tuple $(ip1,ip2)$. In order, the $\mathtt{EventuallyClosed}$ rule states that the connection between IP addresses $i1$ and $i2$ must end eventually within time $d$; finally, the monitor rule $\mathtt{BufferSafe}$ states that it is always the case that if there is an event of $binary$ access in the network log then eventually within time $100$ there must be a matching event in the webserver log that denotes the closing of the connection.

\textbf{Password guessing attacks.}
Assuming the system accepts telnet connections to remote login; it won't take long until the password is compromised if an attacker is permitted to guess any amount of passwords for a specific user name. If a user guesses an incorrect password more than three times in a short amount of time, most systems end the telnet session. To prevent an attacker from succeeding by starting numerous temporary sessions, some systems cap the total number of unsuccessful login attempts over a longer time frame; in this case we need to have access to the host's audit logs. The formulas that capture this scenario are the following:
\begin{align*}
    & \underline{\text{max}}\ \mathtt{Failure}() = (type=login)\land \neg success\\
    & \underline{\text{max}}\ \mathtt{Guess}(\underline{\text{string}}\ i) = (ip=i)\land \mathtt{Failure}()\\
    & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i,\underline{\text{int}}\ C) = (time-t<d)\\
    & \quad\quad\rightarrow((\mathtt{Guess}(i)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i,C)))  \\
    & \quad\quad\land(\neg \mathtt{Guess}(i)\rightarrow\mathtt{Counter}(t,d,c,i,C))) \\
    & \underline{\text{mon}}\ \mathtt{PassGuessSafe} =\ \mathtt{Always}(\mathtt{Failure}()\rightarrow\mathtt{Counter}(time,300,1,ip,3)) 
\end{align*}
as we can see, the rules here are numerous. In $\mathtt{Counter}$, the arguments $t, d, c, i, C$ represent the rule invocation time, the timeout period, the current number of unsuccessful-guesses count, the source IP address doing the guess, and the threshold count. When the number of guesses from the IP address $i$ exceeds the threshold count $C$ within the timeout period $d$, an attack has taken place. The parameterized rule $\mathtt{Counter}$ starts with the initial count $c$ set to 1 whenever there is a $\mathtt{Failure}$ during login. The number of guesses is increased by one for each instance of an event signifying a failed login attempt from the same IP address within the timeout period $d$. The rule $\mathtt{Counter}$ also determines whether the local counter $c$ surpasses the threshold $C$ within time d, and if it does, the entire rule become false, indicating an attack. Finally, with the rule $\mathtt{PassGuessSafe}$ we assure that whenever there is a failed login from a specific IP address, we make sure that within time $300$ seconds the number of login-failures from the same IP address must be less than or equal to $3$.

\textbf{Port Scanning attacks.}
According to~\cite{kanlayasiri2000rule}, Port scanning attack is a method for discovering exploitable communication channels that has been used for a long time. The key idea is to probe the network ports and then keep the information about them that are useful for an attack.
Hackers frequently employ a port scan approach to find gaps or weak spots in a network. Attackers can use a port scan attack to identify open ports and determine if they are accepting or rejecting data; additionally, it can show whether a company uses firewalls or other active security measures. 
Assuming that our server has multiple network inferfaces, thus multiple IP addresses; the victim machine can presume that port scans are malicious when the counter of reached ports overcomes a specific threshold in a short amount of time. The formulas are the following:
\begin{align*}
    & \underline{\text{max}}\ \mathtt{NewPort}(\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S) = (i1=ip1)\land(i2=ip2)\land (port \notin S)\\
    & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S,\underline{\text{int}}\ C) = (time-t<d)\\
    & \quad\quad\rightarrow((\mathtt{NewPort}(i1,i2,S)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i1,i2,S\cup\{port\},C)))  \\
    & \quad\quad\land(\neg \mathtt{NewPort}(i1,i2,S)\rightarrow\mathtt{Counter}(t,d,c,i1,i2,S,C))) \\
    & \underline{\text{mon}}\ \mathtt{PortScanSafe} =\ \mathtt{Always}(\mathtt{Counter}(time,100,1,ip1,ip2,\{port\},10)) 
\end{align*}
in this case we have two IPs; $ip1$ is the source IP, while $ip2$ is the destination one. The parameterized $\mathtt{Counter}$ rule states that the number of port scans between a source and destination IP  address pair $(i1,i2)$ recorded do not surpass a threshold $C$ within time $d$. The rule $\mathtt{NewPort}$ checks if the port number,
involved in any communication between the IP addresses $i1$ and $i2$, exists in the set $S$ of all port numbers (involved in all communications between $i1$ and $i2$) within the timeout period $d$.

\section{Conclusions}
In conclusion, we have introduced a new temporal logic named EAGLE, which serves as the foundation for the MONID framework, which is a useful tool for system intrusion detection.
MONID continuously monitor these attack specifications during system execution and when the observed execution contradicts the monitored formula, we detect an intrusion. We have showcased this by creating formulas to identify several common attacks and testing our monitoring algorithm on large event logs provided by DARPA. These examples can serve as templates for specifying many other types of attacks. The obtained results were excellent, both in terms of detected attacks and system overhead.

>>>>

\section{MONID: Structure analysis}
MONID is a prototype that can detect intrusions either online or offline. System-level event data is delivered to MONID server either directly from relevant log files (offline) or online via application code that has been properly instrumented. In the online mode, MONID runs as a server that receives streams of events from various sources and to generate these events, the different logging modules are instrumented so that they filter and send relevant events to the server.
To create a single event trace, the server combines events from numerous sources according to timestamps and preprocesses them into an abstract intermediate form. Note that sometimes it is necessary to watch events from multiple sources in order to detect certain attacks; then, the MONID monitor checks this event trail against a predetermined criteria, and if the specification is broken, an intrusion alarm is raised. While in offline mode, MONID reads various log files and sends an event corresponding to each log entry to the server. The server then processes the event stream as before to detect an intrusion.

@@

Some evaluations were also done in~\cite{naldurg2004temporal}, using the DARPA Intrusion Detection Evaluation dataset~\cite{mchugh2000testing}. The experiment was done using offline mode against some of the attacks presented in Section~\ref{attacks}. More specifically, against the smurf, port-sweep and password-guessing attacks, the experiments detected 5 password-guess attack and 2 port-sweep attack in the logs and the performance overheads for monitoring the Port-Sweep and Password-Attacks are reported in Figure~\ref{graph1} and~\ref{graph2}, respectively. The X-axis in both graphs shows the number of events we are monitoring, as they are obtained from our logs. Each data point is the average overhead calculated for intervals of 10 and 1000 events respectively. The Y-axis plots the ratio between the time spent by the monitor vs. the time between the generation of the events in the actual log. As long as this ratio is less than 1, our monitoring is feasible. The results show that the average overhead is around $0.009$ for Port-Sweep attacks and $0.016$ for Password-Guessing attacks.
Although the used hardware is now outdated and obsolete, the overhead was rather modest, thus this method should still be theoretically helpful today.

@@

\section{Example of attacks}\label{attacks}
In this section, we give some examples of how EAGLE can be utilized to build formulas that correlate to the desired execution trace characteristics of a system under observation. In this context, a trace $\sigma$ that violates this specification is referred to as an intrusion or an attack; to explain how the framework may be used, we use examples from actual attacks illustrated in~\cite{naldurg2004temporal}. These illustrations demonstrate the expressive power of EAGLE formalism as well as its numerous features. In all the following examples, the rule Always is the one defined in Section~\ref{sec:ltl}.

\textbf{Smurf attacks.}
A smurf attack~\cite{zargar2009identification} is a type of Distributed Denial of Service (DDoS) attack that targets computer networks; in this scenario, the attacker sends a large volume of Internet Control Message Protocol (ICMP) echo request packets (commonly known as "ping" packets) to a network's broadcast address. This broadcast address is the one that reaches all devices on the targeted network. The attacker spoofs the source IP address of these ICMP packets to make it appear as if they are originating from the victim's IP address and since the broadcast address sends the request to all devices on the network, all devices will respond with ICMP echo replies to the victim's IP address. This results in a flood of traffic overwhelming the victim's network and causing it to become unreachable or severely degraded in performance. To detect this kind of attack, we need to work on network logs; we will use a network tool called \textit{tcpdump}. The monitored formula that describes the absence of the attack is the following:
\begin{align*}
    & \underline{\text{max}}\ \mathtt{SmurfAttack()} = (type = \text{"ICMP"})\land isBroadcast(ip) \\
    & \underline{\text{mon}}\ \mathtt{SmurfSafety} =\ \mathtt{Always}(\neg\ \mathtt{SmurfAttack()}) 
\end{align*}
as we can see, the \textit{type} field of a record schema describes the type of network packet, while the \textit{ip} field is the destination of the packet. The $\mathtt{SmurfAttack}$ rule determines whether the packet is a ping packet and whether the source's IP address is a broadcast ip, i.e. all the host bits are set to 1. Thus, \textit{"Always there is no attack"} can be used to describe how the system behaves well in response to this attack.

\textbf{Cookie-stealing attacks.}
A cookie is a technique for tracking sessions that a web application server sends to a web client and uses to keep track of client-specific session data. These cookies that serve as authentication tokens are automatically included in requests by clients and in this scenario we assume that a session is exclusively identified by the client's IP address. When a malicious user uses an old cookie provided by a different IP address to take over a session, it is considered an attack. In order to monitor this attack, we need to look at a web-server (application-level) log that contains a record of all sessions that the server participates in, along with session-specific state information.
According to the formula below, a specific cookie must always be used by the same IP address.
\begin{align*}
    & \underline{\text{min}}\ \mathtt{Hijack}(\underline{\text{string}}\ c,\underline{\text{string}}\ i) = ((name=c)\land\neg(ip=i))\lor\odot\ \mathtt{Hijack}(c,i) \\
    & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\neg\mathtt{Hijack}(name,ip)) 
\end{align*}
A trace that violates this formula therefore encodes a cookie-stealing attack. The rule $\mathtt{Hijack}$ checks whether it occurs a situation in which we have the same cookie $c$ associated with two different ip $i$; if it's true, then it means that a cookie has been stolen and reused by an attacker. With $\mathtt{CookieSafe}$ we want to check that it never occurs a situation just like the one described.

\textbf{Multi-domain Buffer Overflows attacks.}
According to~\cite{piromsopa2006buffer}, a buffer-overflow attack is an attack that (possibly implicitly) uses memory-manipulating operations to overflow a buffer which results in the modification of an address to point to malicious or unexpected code. In this case, we will integrate data from various log sources; to be more exact, we will examine both network and web server access logs to determine when a server buffer overflow attack has been performed. We search through network packets for binary data; then the web server's access logs are examined to see if a corresponding event, in which the web server correctly terminates the connection after receiving some binary data, can be located. If no matching log record is discovered within a predetermined timeout, it means that the buffer overflow attack was successful. The following formulas capture this scenario:
\begin{align*} 
    & \underline{\text{min}}\ \mathtt{EventuallyClosed}(\underline{\text{long}}\ t,\underline{\text{long}}\ d,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2) = (time-t<d)\\
    & \quad\quad\land ((ip1=i1\land ip2=i2\land log=web\land type=closed)\lor \bigcirc\mathtt{EventuallyClosed}(t,d,i1,i2) ) \\
    & \underline{\text{mon}}\ \mathtt{BufferSafe} =\ \mathtt{Always}((log=network\land type=binary) \\
    & \quad\quad\rightarrow\mathtt{EventuallyClosed}(time,100,ip1,ip2))
\end{align*}
In each record we have the $log$ field which indicates the source of the log, the $time$ represents the instant in which the event occurs, $ip1$ represents the source address, $ip2$ represents the destination IP address and finally the $type$ field can be $binary, closed,\ldots$ We assume that a connection is identified by the tuple $(ip1,ip2)$. In order, the $\mathtt{EventuallyClosed}$ rule states that the connection between IP addresses $i1$ and $i2$ must end eventually within time $d$; finally, the monitor rule $\mathtt{BufferSafe}$ states that it is always the case that if there is an event of $binary$ access in the network log then eventually within time $100$ there must be a matching event in the webserver log that denotes the closing of the connection.

\textbf{Password guessing attacks.}
Assuming the system accepts telnet connections to remote login; it won't take long until the password is compromised if an attacker is permitted to guess any number of passwords for a specific user name. If a user guesses an incorrect password more than three times in a short amount of time, most systems end the telnet session. To prevent an attacker from succeeding by starting numerous temporary sessions, some systems cap the total number of unsuccessful login attempts over a longer time frame; in this case we need to have access to the host's audit logs. The formulas that capture this scenario are the following:
\begin{align*}
    & \underline{\text{max}}\ \mathtt{Failure}() = (type=login)\land \neg success\\
    & \underline{\text{max}}\ \mathtt{Guess}(\underline{\text{string}}\ i) = (ip=i)\land \mathtt{Failure}()\\
    & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i,\underline{\text{int}}\ C) = (time-t<d)\\
    & \quad\quad\rightarrow((\mathtt{Guess}(i)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i,C)))  \\
    & \quad\quad\land(\neg \mathtt{Guess}(i)\rightarrow\mathtt{Counter}(t,d,c,i,C))) \\
    & \underline{\text{mon}}\ \mathtt{PassGuessSafe} =\ \mathtt{Always}(\mathtt{Failure}()\rightarrow\mathtt{Counter}(time,300,1,ip,3)) 
\end{align*}
as we can see, the rules here are numerous. In $\mathtt{Counter}$, the arguments $t, d, c, i, C$ represent the rule invocation time, the timeout period, the current number of unsuccessful-guesses count, the source IP address doing the guess, and the threshold count. When the number of guesses from the IP address $i$ exceeds the threshold count $C$ within the timeout period $d$, an attack has taken place. The parameterized rule $\mathtt{Counter}$ starts with the initial count $c$ set to 1 whenever there is a $\mathtt{Failure}$ during login. The number of guesses is increased by one for each instance of an event signifying a failed login attempt from the same IP address within the timeout period $d$. The rule $\mathtt{Counter}$ also determines whether the local counter $c$ surpasses the threshold $C$ within time d, and if it does, the entire rule becomes false, indicating an attack. Finally, with the rule $\mathtt{PassGuessSafe}$ we assure that whenever there is a failed login from a specific IP address, we make sure that within $300$ seconds the number of login-failures from the same IP address must be less than or equal to $3$.

\textbf{Port Scanning attacks.}
According to~\cite{kanlayasiri2000rule}, Port scanning attack is a method for discovering exploitable communication channels that has been used for a long time. The key idea is to probe the network ports and then keep the information about them that is useful for an attack.
Hackers frequently employ a port scan approach to find gaps or weak spots in a network. Attackers can use a port scan attack to identify open ports and determine if they are accepting or rejecting data; additionally, it can show whether a company uses firewalls or other active security measures. 
Assuming that our server has multiple network interfaces, thus multiple IP addresses; the victim machine can presume that port scans are malicious when the counter of reached ports overcomes a specific threshold in a short amount of time. The formulas are the following:
\begin{align*}
    & \underline{\text{max}}\ \mathtt{NewPort}(\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S) = (i1=ip1)\land(i2=ip2)\land (port \notin S)\\
    & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S,\underline{\text{int}}\ C) = (time-t<d)\\
    & \quad\quad\rightarrow((\mathtt{NewPort}(i1,i2,S)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i1,i2,S\cup\{port\},C)))  \\
    & \quad\quad\land(\neg \mathtt{NewPort}(i1,i2,S)\rightarrow\mathtt{Counter}(t,d,c,i1,i2,S,C))) \\
    & \underline{\text{mon}}\ \mathtt{PortScanSafe} =\ \mathtt{Always}(\mathtt{Counter}(time,100,1,ip1,ip2,\{port\},10)) 
\end{align*}
in this case we have two IPs; $ip1$ is the source IP, while $ip2$ is the destination one. The parameterized $\mathtt{Counter}$ rule states that the number of port scans between a source and destination IP  address pair $(i1,i2)$ recorded does not surpass a threshold $C$ within time $d$. The rule $\mathtt{NewPort}$ checks if the port number involved in any communication between the IP addresses $i1$ and $i2$ exists in the set $S$ of all port numbers (involved in all communications between $i1$ and $i2$) within the timeout period $d$.

\section{Conclusions}
In conclusion, we have introduced a new temporal logic named EAGLE, which serves as the foundation for the MONID framework, which is a useful tool for system intrusion detection.
MONID continuously monitors these attack specifications during system execution and when the observed execution contradicts the monitored formula, we detect an intrusion. We have showcased this by creating formulas to identify several common attacks and testing our monitoring algorithm on large event logs provided by DARPA. These examples can serve as templates for specifying many other types of attacks. The obtained results were excellent, both in terms of detected attacks and system overhead.
